---
title: "DPs Syntax in acquisition"
subtitle: "A case study on Italian L2 by Czech ad Slovak learners"
author: "Marco Petolicchio"
institution: "Dept. of Romance Languages. Palacky University, Olomouc (CZ)"
abstract: "Czech and Slovak are languages which don't exhibit a manifest position for the Articles in the Determiner Phrase. The aim of this paper is to show how this structure is accessed during the learning of Italian, a language which presents the articles as for the standard behavior for nouns.\\par
	\\textbf{Keywords:} Determiner Phrase, Italian L2, Second Language Acquisition, Syntax"
site: bookdown::bookdown_site
documentclass: scrartcl
output:
  bookdown::gitbook: 
    split_by: none
    split_bibliography: false
    config:
      sharing: false
  bookdown::pdf_book:
    template: null
    pandoc_args:
    includes: 
      in_header: latex/preamble.tex
    latex_engine: xelatex
    citation_package: natbib
    keep_tex: yes
classoption: [a4paper,twoside,12pt,chapterprefix=false,listof=flat]
bibliography: [bibliography.bib]
biblio-style: apa
link-citations: yes
github-repo: 
description: ""
always_allow_html: yes
---


```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(knitr)

options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
df <- read.csv('data/sentences.csv', header = T, stringsAsFactors = F)

```


\clearpage





# Introduction


Czech (`ces`) and Slovak (`slk`) are languages of the Slavic branch in the Indo-european family. Alongside a certain morphological complexity in noun declension systems, these languages --except for Bulgarian (\@ref(exm:dpBul)) and Macedonian [@wals-37]-- don't show an overt realization of the Determiner position inside the noun phrase (\@ref(exm:dpNo)) [@harkins1953]. 
Conversly, Italian (`ita`) and the other romance languages explicit that position as a default behaviour, usually with a free morpheme preceding the noun (\@ref(exm:dpPre)) or by cliticization of the definite article (\@ref(exm:dpRon)):



(@) Articleless (\#exm:dpNo)  
	a. 	`ces` [@veselovska2014 14]  
		\gll 	Chlapec/Marie/Ona/každý	miluje 		ryby/ \{své 	rodiče\}.  
				Boy/Marie/She/Everyone 	Love.3sg 	Fish/ \{POSS 	parent\}  
		\glt 	"SUBJ loves [the] fish/ his parents"
	a. 	`slk` [@krizomkrazomA1 113]  
		\gll 	Večer 	čítam 		knihy, 	píšem 		referatý...  
				Evening	Read.1SG 	Book.PL Write.1SG	Paper.PL  
		\glt 	"In the evening I read [the] books, I write (school) papers ..."

(@) Proclitic (\#exm:dpPre)  
	a.  `ita` [@bianco2017 60]  
		\gll 	Il 		terremoto 	ha 		distrutto			la 			città.  
			 	ART.DEF Earthquake 	AUX.3sg Destroy.PTCP.PST 	ART.DEF 	City  
		\glt  	"The earthquake destroyed the city"
	a. 	`fro` [@dedole2008 3261]  
		\gll 	La 			dame 	estoit 	devant 	la 			sale.  
				ART.DEF 	Girl 	Be.3sg 	ADV		ART.DEF 	Room  
		\glt 	"The *dame* was in front to the room"

(@) Enclitic
	a. `ron` [@cojocaru2003 45]  (\#exm:dpRon)  
		\gll 	Prieten=ul 			meu 	este 		aici.  
				Friend=ART.DEF 		POSS	Be.PRES.3sg Here  
		\glt  	"My friend is here"
	a. `bul` [@leafgren2011 37]  (\#exm:dpBul)  
		\gll 	Къде 	е 		книга=та 		ми?  
				Where 	Be.3SG 	Book=ART.DEF 	POSS.1SG  
		\glt 	"Where is my book?"


The general idea of this paper is to address the question of how linguistic structures which are not overtly marked in L1 can be accessed during the acquisition of a target language that show them. While doing this can be either both purely speculative as grounded on actual data, I will show how the usage of a target collection of linguistic corpora can be useful to test the main hypotheses into narrower facts. The language under observation are indeed a few: on one side `ces` and `slk` as native languages--with no overt position for the articles--on the other `ita` as target language.

The section \@ref(sec:theoryBg) provides a theoretical discussion on the top of different theories inside the Generative framework [@chomsky1995] on the status of DP and NP. The section \@ref(sec:caseStudy) is twofold: firstly I present the methods used into the current analysis in terms of *reproducibility* of the research, the policies of data-collection and an analysis of the expected results; while the second subsection is built upon a case study made off to test some hypotheses about the categorial differences of DPs during the acquisition of `ita` by `ces` and `slk` native speakers involved in the test.
A summary conclusion (Section \@ref(sec:concl)) closes the paper.



# Theoretical background {#sec:theoryBg}

One of the main area of research in Generative studies on Second Language Acquisition (`GenSLA`) regards the investigation about how the linguistic structures can be accessed in L2 and how the transitional stages of acquisition work into the learning continuum [@rothmanslabakova2017].
In this paper I wish to focus on a narrow structure, e.g. the NP-shell.

Symmetries amongst the DP/NP phrase and the whole sentence are often ereferenced in terms of structure building and phase-related properties [@chomsky2013; @chomsky2015]


```{r tree1, echo=FALSE, engine='tikz', out.width='100%', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'svg', fig.cap='Structural comparison between NP-over-DP and DP-over-NP Hypotheses', engine.opts = list(template = "latex/tikz2pdf.tex"), fig.align='center'}

	\begin{tikzpicture}
		\begin{scope}
			\Tree [ .NP [ .DP D ] [ .N N [ .{...} ] ] ] 
		\end{scope}
		\begin{scope}[xshift=12em]
			\Tree [.DP D [ .NP N [ .{...} ] ] ]
		\end{scope}
	\end{tikzpicture}

```

Two hypothesis: DP as Spec or Phrase.












## Implications for the DP-Hypothesis in Italian




## Implications for the DP-Hypothesis in Czech









# Case Study {#sec:caseStudy}








## Retrieve the linguistic data



```{r graphCorpora, echo=TRUE, fig.width=7, fig.height=4, message=FALSE, fig.cap='Number of texts by different Corpora'}
ggplot(df, aes(fill=df$L1, y=abs(rnorm(nrow(df))), x=df$Corpus)) + 
	geom_bar( stat="identity") + labs(x="Corpora", y="Number of texts", color="Native Language")
```






## Perform a test

Provide 10 noun phrases with different referential structure in terms of animateness with a neutralized article. The students are invited to write full sentences for each one. 

Goals:

* Test the syntactic distribution of phrases (SUBJ vs OBJ)
* Test the distribution of DETs ([+def], [-def], Ø)










# Conclusion {#sec:concl}












# Financial coverage {-}

This work was partially supported by the funding grant XXX.





# Abbreviations {-}

Languages are indicated by the abbreviations provided in the ISO 639-3 format [@iso639-3]. Morphological glosses styles adher to the widespreadly recognized *Leipzig Glossing Rules* [@leipzigGlossingRules], while other abbreviations respect [@boeckxListOfAbbreviations].



