---
title: "DPs Syntax in acquisition"
subtitle: "A case study on Italian L2 by Czech ad Slovak learners"
author: "Marco Petolicchio"
institution: "Dept. of Romance Languages. Palacky University, Olomouc (CZ)"
abstract: "Czech and Slovak are languages which don't exhibit a manifest position for the Articles in the Determiner Phrase. The aim of this paper is to show how this structure is accessed during the learning of Italian, a language which presents the articles as for the standard behavior for nouns.\\par
	\\textbf{Keywords:} Determiner Phrase, Italian L2, Second Language Acquisition, Syntax"
site: bookdown::bookdown_site
documentclass: scrartcl
output:
  bookdown::gitbook: 
    split_by: none
    split_bibliography: false
    config:
      sharing: false
  bookdown::pdf_book:
    template: null
    pandoc_args:
    includes: 
      in_header: latex/preamble.tex
    latex_engine: xelatex
    citation_package: natbib
    keep_tex: yes
classoption: [a4paper,twoside,12pt,chapterprefix=false,listof=flat]
bibliography: [bibliography.bib]
biblio-style: apa
link-citations: yes
github-repo: 
description: ""
always_allow_html: yes
---


```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(knitr)

options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
df <- read.csv('data/sentences.csv', header = T, stringsAsFactors = F)

```


\clearpage





# Introduction


Czech (`cs`) and Slovak (`sk`) are languages of the Slavic branch in the Indo-european family. Alongside a certain morphological complexity in noun declension systems, these languages -- except for Bulgarian and Macedonian [@wals-37 in @wals] -- don't show an overt realization of the Determiner position inside the noun phrase [@harkins1953]. Conversly, Italian (`it`) and the other romance languages explicit that position as a default behaviour, usually with a free morpheme before the noun^[Otherwise, Romanian uses a intralexematic suffixation for the article, which falls in the end of the word.]:


(@)	Slavic languages	
	a.	(`cs`, *Reference*)  
		\gll 	Viděl 		 jsem 	Ø/?ten 	chlapec který 	běží  
			 	See.PTCP.PST AUX.1sg DEM 	Boy.SG	pro 	Run.3sg  
		\glt 	    
(@) Romance languages
	a.	(`it`, *Reference*)  
		\gll 	Ho 	 	visto 			un/il/\*Ø 	ragazzo che corre  
			 	AUX.1sg See.PTCP.PST 	DET 		Boy.SG 	pro Run.3sg  
		\glt  
	a.	(`fr`, *Reference*)  
		\gll 	J'ai 		vu 			 un/le/?Ø 	garçon 	courir  
				I.AUX.1sg 	See.PTCP.PST DET 		Boy.SG 	Run  
		\glt  









# Theoretical background




Two hypothesis: DP as Spec or Phrase.

```{r tree1, echo=FALSE, engine='tikz', out.width='70%', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'svg', fig.cap='DP and NP phrases', engine.opts = list(template = "latex/tikz2pdf.tex"), fig.align='center'}

\begin{tikzpicture}
	\begin{scope}
		\Tree [ .NP [ .DP D ] [ .N N [ .{...} ] ] ] 
	\end{scope}
	\begin{scope}[xshift=12em]
		\Tree [.root(X) [ .mod(X) head(mod(X)) ] [ .proj(X) head(X) [ .comp(X) ] ] ]
	\end{scope}
	\begin{scope}[yshift=12em]
		\Tree [.DP D [ .NP N [ .{...} ] ] ]
	\end{scope}
	\begin{scope}[yshift=12em, xshift=12em]
		\Tree [.root(Y) head(Y) [ .root(X) head(X) [ .comp(X) ] ] ]
	\end{scope}
\end{tikzpicture}
```






## Determiners in Italian








## Implications for the DP-Hypothesis in Italian









# Case Study








## Retrieve the linguistic data 



```{r graphCorpora, echo=TRUE, fig.width=7, fig.height=4, message=FALSE, fig.cap='Number of texts by different Corpora'}
ggplot(df, aes(fill=df$L1, y=abs(rnorm(nrow(df))), x=df$Corpus)) + 
	geom_bar( stat="identity") + labs(x="Corpora", y="Number of texts", color="Native Language")
```



## Perform a test 

Provide 10 noun phrases with different referential structure in terms of animateness with a neutralized article. The students are invited to write full sentences for each one. 

Goals:

* Test the syntactic distribution of phrases (SUBJ vs OBJ)
* Test the distribution of DETs ([+def], [-def], Ø)










# Conclusion












# Financial coverage

This work was partially supported by the funding grant XXX.





# Abbreviations

Languages are indicated by the abbreviations provided in the standard ISO 639-1 format [@iso639-1]. The linguistic abbreviations adher to the widespreadly recognized *Leipzig Glossing Rules* [@leipzigGlossingRules].





# References